---
title: "DESeq ES cells loop analysis"
output: html_notebook
---


Importing the data with and without DESeq

```{r}
library(dplyr)


# df_es_loop_deseq <- read.csv("Z:/Nathan/Resultats/HRS-seq-SOURIS/HRS-seq-SOURIS-mm9-avec-DEseq/listHRS_ES_avecDESeq_refboo.csv",
#                         sep = ";")

df_es_loop_deseq <- read.csv("C:/local_data/HRS-seq-SOURIS/HRS-seq-SOURIS-mm9-avec-DEseq/listHRS_ES_avecDESeq_refboo.csv",
                        sep = ";")



df_es_loop_deseq <- select(df_es_loop_deseq, 
                      c(chromosome, deb, fin, pvalFDR_DEseq, FC))

colnames(df_es_loop_deseq) <- c("chromosome", "deb", "fin", "pval_DEseq", "FC_DEseq")




# 
# df_es_loop_edger <- read.csv("Z:/Nathan/Resultats/HRS-seq-SOURIS/HRS-seq-SOURIS-mm9-sansDEseq/listHRS_ES_BOO.csv",
#                         sep = ";")


df_es_loop_edger <- read.csv("C:/local_data/HRS-seq-SOURIS/HRS-seq-SOURIS-mm9-sansDEseq/listHRS_ES_BOO.csv",
                        sep = ";")


df_es_loop_edger <- select(df_es_loop_edger, 
                      c(chromosome, deb, fin, pvalFDR_edgeR_exact, FC))

colnames(df_es_loop_edger) <- c("chromosome", "deb", "fin", "pval_edgeR", "FC_edgeR")


df_es_loop_combined <- bind_rows(df_es_loop_deseq, df_es_loop_edger)


```

```{r}
library(ggplot2)
library(ggpointdensity)
library(viridis)
library(ggExtra)


volcano_DEseq_loop <- ggplot(data = df_es_loop_deseq,
                        mapping = aes(y = -log10(pval_DEseq),
                                      x = log2(FC_DEseq))) +
                 geom_point() +
                 geom_pointdensity() +
                 scale_color_viridis() +
                 scale_x_continuous(limits = c(0, 10))+
                 ylim(c(0,40))

volcano_DEseq_loop



volcano_edger_loop <- ggplot(data = df_es_loop_edger, 
                        mapping = aes(y = -log10(pval_edgeR),
                                      x = log2(FC_edgeR))) +
                 geom_point() +
                 geom_pointdensity() +
                 scale_color_viridis() +
                 scale_x_continuous(limits = c(0, 10)) +
                 ylim(c(0,40))
   

volcano_edger_loop       


```



Creating a dataframe combining the DEseq and edgeR fragments in 

```{r}
df_es_loop_fused_na <- df_es_loop_combined

df_es_loop_fused_na[is.na(df_es_loop_fused_na)] <- 0


df_es_loop_fused <- df_es_loop_fused_na %>% 
      
      group_by(chromosome, deb, fin) %>% 
      
      summarise(across(c(pval_DEseq, 
                         FC_DEseq, 
                         pval_edgeR, 
                         FC_edgeR), 
                       sum))
```





```{r}
scatter_FC_loop <- ggplot(data = df_es_loop_fused,
                     mapping = aes(x = FC_DEseq,
                                   y = FC_edgeR)) +
              scale_x_continuous(trans = "log2") +
              scale_y_continuous(trans = "log2") +
              geom_pointdensity() +
              scale_color_viridis()

scatter_FC_loop



```




```{r}
scatter_pval_loop <- ggplot(data = df_es_loop_fused,
                       mapping = aes(x = -log10(pval_DEseq),
                                     y = -log10(pval_edgeR))) +
              geom_pointdensity() +
              scale_color_viridis() +
              scale_x_continuous(limits = c(0, 40)) +
              scale_y_continuous(limits = c(0, 40)) +
              geom_abline(slope = 1, intercept = 0)
   

ggMarginal(scatter_pval_loop, type = "histogram")
```



Extraction of the common fragments and specific fragments

```{r}

df_es_loop_common_only <- filter(df_es_loop_fused, 
                           (pval_DEseq != 0))

df_es_loop_edger_only <- filter(df_es_loop_fused,
                           (pval_DEseq == 0))



```


Volcano plot of the edge R exclusive fragments and common fragments

```{r}

volcano_edger_only_loop <- ggplot(data = df_es_loop_edger_only,
                                  mapping = aes(x = log2(FC_edgeR),
                                                y = -log10(pval_edgeR))) +
                           geom_pointdensity() +
                           labs(title = "fragments excluded by DESeq") +
                           scale_x_continuous(limits = c(0, 10))+
                           scale_y_continuous(limits = c(0, 40))+
                           scale_color_viridis()



ggMarginal(volcano_edger_only_loop, type = "histogram")



volcano_common_only_loop <- ggplot(data = df_es_loop_common_only,
                                  mapping = aes(x = log2(FC_edgeR),
                                                y = -log10(pval_edgeR))) +
                            geom_pointdensity() +
                            labs(title = "common fragments between DESeq and edgeR") +
                            scale_x_continuous(limits = c(0, 10))+
                            scale_y_continuous(limits = c(0, 40))+
                            scale_color_viridis()



ggMarginal(volcano_common_only_loop, type = "histogram")





```


Checking the representation of chromosomes in the top DESeq excluded fragments

```{r}
df_es_loop_edger_only$count <- 1 

df_representation_loop <- df_es_loop_edger_only %>%
   
   group_by(chromosome) %>%
   
   summarise(across(count, sum))


df_representation_loop$ratio <- df_representation_loop$count /
                                sum(df_representation_loop$count)



df_es_loop_fused$count_global <- 1




df_counts_global_loop <- df_es_loop_fused %>% 
   group_by(chromosome) %>% 
   summarise(across(count_global, sum))

df_counts_global_loop$ratio_global <- df_counts_global_loop$count_global /
                                 sum(df_counts_global_loop$count_global)


df_es_counts_global_loop <- arrange(df_counts_global_loop, chromosome)

df_representation_loop <- arrange(df_representation_loop, chromosome)



df_representation_loop$ratio_global <- df_counts_global_loop$ratio_global

df_representation_loop$count_global <- df_counts_global_loop$count_global
```



Plotting the fragment's ratios

```{r}
ggplot(df_representation_loop,
       mapping = aes(x = "", y = ratio*100, fill = chromosome)) +
   geom_bar(stat = "identity", width = 1, colour = "black") +
   labs(title = "edgeR only") +
   coord_polar("y", start=0)
   

ggplot(df_representation_loop,
       mapping = aes(x = "", y = ratio_global*100, fill = chromosome)) +
   geom_bar(stat = "identity", width = 1, colour = "black") +
   labs(title = "all") +
   coord_polar("y", start=0)


```


1D clustering to find regions standing out


```{r}
library(Ckmeans.1d.dp)
library(RColorBrewer)


df_es_loop_edger_only$center <- df_es_loop_edger_only$deb + 
                                (df_es_loop_edger_only$fin - 
                                 df_es_loop_edger_only$deb)/2

first_group = function(x) x %>%
    select(group_cols()) %>%
    distinct %>%
    ungroup %>%
    slice(1) %>%
    { semi_join(x, .)}





chroms <- group_by(df_es_loop_edger_only, chromosome)

isolate_group = function(df, n) df %>%
    select(group_cols()) %>%
    distinct %>%
    ungroup %>%
    slice(n) %>%
    { semi_join(x, .)}

test <- first_group(chroms)



results <- (Ckmeans.1d.dp(test$center))[cluster]

clustering_loop <- df_es_loop_edger_only %>% 
                   group_by(chromosome) %>% 
                   mutate(cluster = (Ckmeans.1d.dp(center, c(1, 1000)))$cluster)


```

```{r}

                                  mapping = aes(x = log2(FC_edgeR),
                                                y = -log10(pval_edgeR),
                                                color = cluster)) +
                           geom_point()


scatter_clustering_loop


scatter_clustering_loop <- clustering_loop %>% 
                           group_by(chromosome) %>%
                           ggplot(mapping = aes(x = log2(FC_edgeR),
                                                y = -log10(pval_edgeR),
                                                color = cluster)) +
                           geom_point()



                           

scatter_clustering_loop + facet_wrap()
```


